The Influence of Connectivity on Dung Beetle Communities

### 2025-07-18 corridor_docs/eric_ms_thesis/MS_thesis.Rmd
post-defense edits

All dung beetles were counted and identified to species using @nemes_illustrated_2015 and @edmonds_taxonomic_2023. Fifteen individuals of each species with adequate captures were dried to equilibrium and weighed for biomass measurements. Voucher specimens for each species will be deposited at the Florida State Collection of Arthropods upon completion of all analyses.

@@

I used linear mixed-effects models to compare the influence of patch type on both Shannon Diversity and Simpson's indexes, including block as a random effect to account for spatial variation. Across both diversity models, the block-level random effect standard deviation was slightly greater than the residual error, indicating that variation between blocks accounted for a substantial portion of the overall variability.

@@

### Beetle Abundance

@@

### Beetle Richness and Diversity

@@

@@

Although richness and diversity were the same among treatments, there was notable variability between sampling blocks. Block 8 generally had the lowest species richness and biodiversity while 53n had the highest. While the design of the experimental blocks control for as much as possible, landscape features across the sight are different enough to influence species composition. Several factors may explain this pattern, at the time this study was conducted block 8 had the densest matrix of any of the other sampling blocks which could have hindered the scent permeability of the bait. Other habitat characteristics might have been influential as well, including soil quality and forest cover which can determine where beetles can reproduce [@arellano_response_2008; @conover_phenology_2019] 
Not only do these impact dung beetle movement, but also these landscape features can affect mammal movement, which in turn limits dung availability for beetles. In this regard it is also important to consider other geographical differences, block 8 was nearest to roadways used by all employees at SRS and the eastern side of the sight had large bodies of water. Both of which would effect the land-use by larger mammals [@WOS:000241362400032; @dechen_quinn_landscape_2013; @WOS:000651446900001]. Additionally, deer and feral hogs are the most dominant mammals at the site, two species that do well in patchy landscapes [@castillo-contreras_urban_2018; @fraser_connectivity_2019]. This might explain the lack of variation in richness or diversity in patches while differences are more apparent at greater distances across the site.

@@

<!-- Paragraph 5: Our results largely echo/constrast with previous ones from the SRS and elsewhere  -->

@@

<!-- Paragraph 6: Caveats -->

>>>>

All dung beetles were counted and identified to species using @nemes_illustrated_2015 and @edmonds_taxonomic_2023. Fifteen individuals of each species with adequate captures were dried to equilibrium and weighed for biomass measurements. **[EDIT SENTENCE ABOVE AS FOLLOWS:** Fifteen individuals of each species with adequate captures (was there a numeric cutoff?) were dried at XXX degrees until they reached a stable mass. We weighed each beetle using a BRAND, MODEL microbalance, except for species whose individuals were extremely small. Individual biomass for these species was estimated by weighing batches of N = 5  (i.e., ---------) or N = 3  (i.e., ---------) beetles and then calculating the average biomass per beetle. These values of individual biomass were then used to estimate the total dung beetle biomass -- all species combined and for each individual species - in each block and patch. **]** Voucher specimens for each species will be deposited at the Florida State Collection of Arthropods upon completion of all analyses.

@@

I used linear mixed-effects models to compare the influence of patch type on both Shannon Diversity and Simpson's indexes, including block as a random effect to account for spatial variation. Across both diversity models, the block-level random effect standard deviation was slightly greater than the residual error, indicating that variation between blocks accounted for a substantial portion of the overall variability.  **[need to add to the last paragraph:** that you did these analyses for the dung beetle biomass as well. And it would be cool for guild...**]**

@@

### Dung Beetle Abundance

@@

### Dung Beetle Richness and Diversity

@@

### Dung Beetle Biomass

**ADD RESULTS FOR DB BIOMASS ANALYSES**

@@

Although richness and diversity were the same among treatments, there was notable variability between sampling blocks. Block 8 generally had the lowest species richness and biodiversity while 53n had the highest. While the design of the experimental blocks control for as much as possible, landscape features across the sight are different enough to influence species composition. Several factors may explain this pattern, at the time this study was conducted block 8 had the densest matrix of any of the other sampling blocks which could have hindered the scent permeability of the bait. Other habitat characteristics might have been influential as well, including soil quality and forest cover which can determine where beetles can reproduce [@arellano_response_2008; @conover_phenology_2019] 
Not only do these impact dung beetle movement, but also these landscape features can affect mammal movement, which in turn limits dung availability for beetles. In this regard it is also important to consider other geographical differences, block 8 was nearest to roadways used by all employees at SRS and the eastern side of the sight had large bodies of water. Both of which would effect the land-use by larger mammals [@WOS:000241362400032; @dechen_quinn_landscape_2013; @WOS:000651446900001]. Additionally, deer and feral hogs are the most dominant mammals at the site, two species that do well in patchy landscapes [@castillo-contreras_urban_2018; @fraser_connectivity_2019]. This might explain the lack of variation in richness or diversity in patches while differences are more apparent at greater distances across the site.  **[need to add and discuss soil compaction at 8 here in this paragraph: 1-2 sentences]**

@@

<!-- Paragraph 5: Beetle Biomass  -->

**[Paragraph 6: biomass]** The patterns in dung beetle biomass largeley echo the ones for richness and abundnace,. these have potential implications for ecosystem functioning. PS might want to add someting to the Future directions about "Interestingly, dung beetles appear to be an exception to the globally and taxonomically robust rule that abundance  is negatively correlated with individual biomass....

<!-- Paragraph 6: Our results largely echo/constrast with previous ones from the SRS and elsewhere  -->

@@

<!-- Paragraph 7: spatial scale of experimental plots vs. dung detectability  -->

**[Paragraph 7: spatial scale of experimental plots vs. dung detectability]** Finally, it is important to emphasize that while dung beetles are capable of long distance flight and detecting dung at distances of over X meters, the results I observed are lijkely not the result of a mismatch between the spatial scale of the experimental replicates and dung beetle movement. Put another way, some beetles were almost certainly drawn by the dung used in baits from the matrix into the plots. If this were a big effect we'd expect all plots to be similar to the matrix, or at least to each other. the fact that rectangular are lower suggests that...]

<!-- Paragraph 7: Caveats -->
